US President Prediction

### 2024-11-03 paper/paper.qmd
add predition graph

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(ggpubr)
```

@@

cite R, packages, dataa
The data used in this paper is source from FiveThirtyEight ("538presidentialpolls", 2024). T
We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...

@@

# Results {#set-result}
The results of the Bayesian linear regression models for Kamala Harris and Donald Trump are summarized in @tbl-model_summary_H and @tbl-model_summary_T.

```{r}
#| label: tbl-model_summary_H
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
#| warning: false
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables
library(modelsummary)

@@

```
```{r}
#| label: tbl-model_summary_T
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Trump"
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables

@@

# Display the summary table
kable(model_summary, digits = 3)
```
## Overall trend
The intercept of both model is positive, with Harris at 40.47 and Trump slightly lower than Harris at 40.36. The trend of support rate for both candidates with respect of time is positive but Harris's increment (0.10) is larger than Trump's (0.06).

@@

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(lubridate)
library(janitor) # Load the janitor package for clean_names()
```

@@

cite R, packages, data
The data used in this paper is source from FiveThirtyEight ("538presidentialpolls", 2024). T

We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...

@@

# Results {#sec-result}
The results of the Bayesian linear regression models for Kamala Harris and Donald Trump are summarized in @tbl-model_summary_H and @tbl-model_summary_T.

```{r}
#| label: tbl-model_summary_H
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
#| warning: false
#| 
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables
library(modelsummary)

@@

```{r}

#| label: tbl-model_summary_T
#| echo: false
#| tbl-cap: "Summary statistics of the Bayesian model for Trump"
# Load required packages
library(broom.mixed)  # For tidy model summaries from mixed models
library(dplyr)
library(readr)
library(knitr)  # For nicely formatted tables

@@

# Display the summary table
kable(model_summary, digits = 3)
```

## Overall trend
The intercept of both model is positive, with Harris at 40.47 and Trump slightly lower than Harris at 40.36. The trend of support rate for both candidates with respect of time is positive but Harris's increment (0.10) is larger than Trump's (0.06).

@@

## Prediction Visualization

```{r}
#| label: fig-prediction-graph
#| echo: false
#| tbl-cap: "2024 US Presidential Election Prediction Based Baysian Linear Model"

earliest_date = 2022-10-11

# Define the target end date
last_date <- as.Date("2024-11-05")

# Calculate the number of days from earliest_date to last_date
days_to_last_date <- as.numeric(last_date - earliest_date)

# Generate prediction data frame with days_after_earliest up to days_to_last_date
days_range <- data.frame(days_after_earliest = seq(0, days_to_last_date, by = 1))

# Convert days_after_earliest to actual dates by adding them to earliest_date
days_range$end_date <- earliest_date + days_range$days_after_earliest

# Harris predictions
harris_predictions <- predict(model1, newdata = days_range, se.fit = TRUE)
days_range$harris_pct <- harris_predictions$fit
days_range$harris_se <- harris_predictions$se.fit

# Trump predictions
trump_predictions <- predict(model2, newdata = days_range, se.fit = TRUE)
days_range$trump_pct <- trump_predictions$fit
days_range$trump_se <- trump_predictions$se.fit

# Add actual poll data to plot
just_harris_high_quality$Candidate <- "Harris"
just_trump_high_quality$Candidate <- "Trump"
poll_data <- rbind(just_harris_high_quality, just_trump_high_quality)

# Plot
ggplot(data = days_range, aes(x = end_date)) +
  # Harris line and confidence interval
  geom_line(aes(y = harris_pct, color = "Harris"), linewidth = 1) +
  geom_ribbon(aes(ymin = harris_pct - 1.96 * harris_se, ymax = harris_pct + 1.96 * harris_se, fill = "Harris"), alpha = 0.2) +

  # Trump line and confidence interval
  geom_line(aes(y = trump_pct, color = "Trump"), linewidth = 1) +
  geom_ribbon(aes(ymin = trump_pct - 1.96 * trump_se, ymax = trump_pct + 1.96 * trump_se, fill = "Trump"), alpha = 0.2) +

  # Actual poll data points
  geom_point(data = poll_data, aes(x = end_date, y = pct, color = Candidate), alpha = 0.5) +

  # Axis labels and title
  labs(title = "Polling Trends for Harris and Trump",
       x = "Date",
       y = "Polling Percentage (%)",
       color = "Candidate",
       fill = "Candidate") +

  # Custom x-axis to show every 2 months
  scale_x_date(
    date_breaks = "2 months",
    date_labels = "%b",  # Shows only the month (e.g., "Aug", "Oct")
    limits = c(earliest_date, as.Date("2024-11-05"))
  ) +

  # Highlight Nov 5th specifically with a dashed line
  geom_vline(xintercept = as.numeric(as.Date("2024-11-05")), linetype = "dashed", color = "black") +

  # Color scheme
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_fill_manual(values = c("Harris" = "blue", "Trump" = "red")) +

  # Theme for cleaner look
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate labels slightly to 45 degrees
  )
